Association of Mental Health Factors with COVID-19 Test Status

### 2025-05-02 code/analysis-code/analysis.qmd
Part 6 updates

We create estimates and CIs using complex survery design variables, using the multiple imputed data.

```{r}

# Create survey design objects for each transformed dataset
svy_designs <- lapply(data_imputed, function(data) {
  svydesign(ids = ~PSU, strata = ~STRATA, weights = ~SAMPWEIGHT, data = data , nest=TRUE)
})

# Function to get complete estimates (both 0 and 1) stratified by COVID status
get_complete_stratified_estimates <- function(var_name) {
  results <- data.frame(
    Variable = character(),
    COVID_Status = character(),
    Binary_Value = numeric(),
    Estimate = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    stringsAsFactors = FALSE
  )

  # Loop through COVID status (0 and 1)
  for (covid_val in c(0, 1)) {
    covid_label <- ifelse(covid_val == 0, "Negative", "Positive")

    # Calculate proportions for this COVID status
    props <- lapply(svy_designs, function(design) {
      # Create a subset design
      subdesign <- subset(design, COVID == covid_val)
      # Create formula for the variable
      formula <- as.formula(paste0("~factor(", var_name, ")"))
      # Calculate proportions
      svymean(formula, subdesign, na.rm = TRUE)
    })

    # Pool results
    pooled <- MIcombine(props)

    # Extract results for both binary values (if available)
    coef_names <- names(coef(pooled))

    # Handle different potential naming patterns
    for (bin_val in c(0, 1)) {
      # Check different possible coefficient names
      possible_names <- c(
        paste0("factor(", var_name, ")", bin_val),
        as.character(bin_val)
      )

      # Find the matching coefficient name
      coef_idx <- NULL
      for (name in possible_names) {
        if (name %in% coef_names) {
          coef_idx <- which(coef_names == name)
          break
        }
      }

      # If found, extract estimates
      if (!is.null(coef_idx)) {
        est <- coef(pooled)[coef_idx]
        ci <- confint(pooled)[coef_idx, ]

        # Add to results
        results <- rbind(results, data.frame(
          Variable = var_name,
          COVID_Status = covid_label,
          Binary_Value = bin_val,
          Estimate = round(est * 100, 1),
          CI_lower = round(ci[1] * 100, 1),
          CI_upper = round(ci[2] * 100, 1)
        ))
      }
    }
  }

  return(results)
}

# Variables to analyze
variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Get complete stratified estimates for each variable
complete_results <- lapply(variables, get_complete_stratified_estimates)

@@

print(formatted_results)

@@

# Prepare the data by combining Variable and Binary_Label, pivoting, and formatting the results
table1_imputed <- formatted_results %>%
  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),  # Convert Yes to 1 and No to 0
         Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
  pivot_wider(
    names_from = COVID_Status,
    values_from = c(Estimate, CI_lower, CI_upper),
    names_glue = "{COVID_Status}_{.value}"
  ) %>%
  mutate(
    # Scale the estimates and CI by 0.01
    Estimate_Negative = Negative_Estimate * 0.01,
    CI_Negative = sprintf("(%.2f, %.2f)", Negative_CI_lower * 0.01, Negative_CI_upper * 0.01),
    Estimate_Positive = Positive_Estimate * 0.01,
    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01)
  ) %>%
  select(Variable_Label, Estimate_Negative, CI_Negative, Estimate_Positive, CI_Positive) %>%
  gt() %>%
  tab_header(
    title = "Estimated proportions of mental health related factors by COVID test status"
  ) %>%
  tab_spanner(
    label = "COVID Negative",
    columns = c(Estimate_Negative, CI_Negative)
  ) %>%
  tab_spanner(
    label = "COVID Positive",
    columns = c(Estimate_Positive, CI_Positive)
  ) %>%
  cols_label(
    Variable_Label = "Variable",               # Rename Variable_Label to Variable
    Estimate_Negative = "Estimate",            # Rename Estimate_Negative to Estimate
    CI_Negative = "95% CI",                    # Rename CI_Negative to 95% CI
    Estimate_Positive = "Estimate",            # Rename Estimate_Positive to Estimate
    CI_Positive = "95% CI"                     # Rename CI_Positive to 95% CI
  )

>>>>

We create estimates and CIs using complex survery design variables, using the multiple imputed data. This will find estimates for overall, and by COVID-19 test status.
```{r}
# Create survey design objects for each transformed dataset
svy_designs <- lapply(data_imputed, function(data) {
  svydesign(ids = ~PSU, strata = ~STRATA, weights = ~SAMPWEIGHT, data = data , nest=TRUE)
})

get_complete_stratified_estimates <- function(var_name) {
  results <- data.frame(
    Variable = character(),
    COVID_Status = character(),
    Binary_Value = numeric(),
    Estimate = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    stringsAsFactors = FALSE
  )

  # Loop through COVID status: Negative (0), Positive (1), Overall (NA)
  for (covid_val in c(0, 1, NA)) {
    if (is.na(covid_val)) {
      covid_label <- "Overall"
      subset_designs <- svy_designs  # No subsetting
    } else {
      covid_label <- ifelse(covid_val == 0, "Negative", "Positive")
      subset_designs <- lapply(svy_designs, function(design) subset(design, COVID == covid_val))
    }

    # Calculate proportions for this COVID status or overall
    props <- lapply(subset_designs, function(design) {
      formula <- as.formula(paste0("~factor(", var_name, ")"))
      svymean(formula, design, na.rm = TRUE)
    })

    # Pool results
    pooled <- MIcombine(props)
    coef_names <- names(coef(pooled))

    for (bin_val in c(0, 1)) {
      possible_names <- c(
        paste0("factor(", var_name, ")", bin_val),
        as.character(bin_val)
      )
      coef_idx <- NULL
      for (name in possible_names) {
        if (name %in% coef_names) {
          coef_idx <- which(coef_names == name)
          break
        }
      }

      if (!is.null(coef_idx)) {
        est <- coef(pooled)[coef_idx]
        ci <- confint(pooled)[coef_idx, ]

        results <- rbind(results, data.frame(
          Variable = var_name,
          COVID_Status = covid_label,
          Binary_Value = bin_val,
          Estimate = round(est * 100, 1),
          CI_lower = round(ci[1] * 100, 1),
          CI_upper = round(ci[2] * 100, 1)
        ))
      }
    }
  }

  return(results)
}
variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Get complete stratified + overall estimates for each variable
complete_results <- lapply(variables, get_complete_stratified_estimates)

@@

print(formatted_results)
```

@@

table1_imputed <- formatted_results %>%
  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),
         Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
  pivot_wider(
    names_from = COVID_Status,
    values_from = c(Estimate, CI_lower, CI_upper),
    names_glue = "{COVID_Status}_{.value}"
  ) %>%
  mutate(
    Estimate_Negative = Negative_Estimate * 0.01,
    CI_Negative = sprintf("(%.2f, %.2f)", Negative_CI_lower * 0.01, Negative_CI_upper * 0.01),
    Estimate_Positive = Positive_Estimate * 0.01,
    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01),
    Estimate_Overall = Overall_Estimate * 0.01,
    CI_Overall = sprintf("(%.2f, %.2f)", Overall_CI_lower * 0.01, Overall_CI_upper * 0.01)
  ) %>%
  select(Variable_Label, Estimate_Negative, CI_Negative,
         Estimate_Positive, CI_Positive, Estimate_Overall, CI_Overall) %>%
  gt() %>%
  tab_header(
    title = "Estimated proportions of mental health related factors by COVID test status"
  ) %>%
  tab_spanner(
    label = "COVID Negative",
    columns = c(Estimate_Negative, CI_Negative)
  ) %>%
  tab_spanner(
    label = "COVID Positive",
    columns = c(Estimate_Positive, CI_Positive)
  ) %>%
  tab_spanner(
    label = "Overall",
    columns = c(Estimate_Overall, CI_Overall)
  ) %>%
  cols_label(
    Variable_Label = "Variable",
    Estimate_Negative = "Estimate",
    CI_Negative = "95% CI",
    Estimate_Positive = "Estimate",
    CI_Positive = "95% CI",
    Estimate_Overall = "Estimate",
    CI_Overall = "95% CI"
  )

print(table1_imputed)

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

## Conclusions
We found that indications of poor mental health were common in the population, as INSERT HERE. Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping. However, these differences were relatively small.

Predicitve modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.

>>>>

## Conclusions
We found that indications of poor mental health were common in the population, as ######%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%INSERT HERE. YOU NEED TO MAKE THIS TABLE FOR NON SPLIT BY COVID TEST AND WRITE RESULTS IN. ALSO YOU NEED TO PUT IN CITAIONS Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping. However, these differences were relatively small.

Predictive modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.
